--- 
title: "Assignment 1, Task 1"
author: "Dylan Glave"
date: "1/20/2021"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(patchwork)
library(janitor)
library(here)
library(readxl)
library(gghighlight)
library(ggfortify)
library(plotly)
library(lubridate)
library(RColorBrewer)
library(wesanderson)
library(ggtext)
library(patchwork)
```


# Sierra Nevada Amphibians

## Wrangling

Below I am reading in data for Sierra Nevada Amphibians to a raw dataframe, and making the observations a bit more coder-friendly with lower-case writing and dates in ymd format. 

```{r}
sierra_raw <- read_excel("sierra_amphibians.xlsx") %>% 
  clean_names() %>% 
  mutate(across(where(is.character), tolower)) %>% 
  mutate(survey_date = ymd(survey_date))
  

```

I've now changed the dates to ymd dates, brought all observations to lower case, and cleaned_names (which didn't seem to do much after the other fucntions.)

To visualize this data, I'll make a more focused df. 
This will require grouping amphibians by year, and creating a df excluding "EggMass," from the life stages **Rana muscosa**

From there I make a summary df of **Rana muscosa**, grouped by year, life stage, and total observations. 

```{r}
ramu_by_year <- sierra_raw %>% 
 mutate(year = year(survey_date)) %>% 
  filter(amphibian_species == "ramu") %>% 
  filter(amphibian_life_stage != "eggmass") 
##Quick summary df of ramu amphibians, grouped by year and life stage.
ramu_summary <- ramu_by_year %>% 
  group_by(year, amphibian_life_stage) %>% 
  summarize(total = sum(amphibian_number, na.rm = TRUE))



```

Making a basic plot of ramu life stages across all sites, grouped by year. This plot contains 3 facets of **Rana muscosa**, one for each life stage. With that, we can see the relative abundance of adults and subadults, while the tadpole numbers dwarf these subsets. 

```{r}
rana_year_graph <- ggplot(data = ramu_summary, aes(x = year, y = total, fill = amphibian_life_stage, color = amphibian_life_stage)) +
 geom_col(show.legend = FALSE) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle=45)) +
  labs( x = "Year",
        y ="Total Organisms Found", 
        title =  "Rana muscosa Observations \n (Across all sites") +
  facet_wrap("amphibian_life_stage") +
  scale_fill_manual(values=wes_palette(n=3, name="Royal2")) +
  scale_color_manual(values=wes_palette(n=3, name="Royal2")) 

  
 
```

## Task 1, Part 2: 

We'll play around to get the 5 lakes with the highest counts of *Rana muscosa*, and put them into a bar graph at the end. 

To start, I'll only count adult and subadult observations, summarize them by lake ID, and list the top 5 in a summar df. 

```{r}
ramu_by_lake <-  sierra_raw %>% 
  filter(amphibian_species == "ramu") %>% 
  filter(amphibian_life_stage != "eggmass") %>% 
  filter(amphibian_life_stage != "tadpole") %>% 
  group_by(lake_id) %>% 
  summarize(total = sum(amphibian_number)) %>% 
  slice_max(total, n=5) %>% 
  mutate(lake_id = as.factor(paste("Lake", lake_id))) %>% 
mutate(lake_id = fct_reorder(lake_id, total))
  
  
```

Now I have a summary table, across all years, of Adult and Subadult **Rana muscosa**. In the df, I've made lake_id a factor, added the word "Lake" before each ID number, and reordered the lakes. In their new order, the 5 lakes are listed from highest to lowest counts of **Rana muscosa**

From here, we'll make a column graph of these 5 lakes. 

```{r}
rana_lake_graph <- ggplot(data = ramu_by_lake, aes(x = lake_id, y = total)) +
  geom_col(aes(fill = lake_id, color = lake_id), 
           alpha = 0.8,
           show.legend = FALSE) +
  scale_fill_manual(values=wes_palette(n=5, name="Royal2")) +
  scale_color_manual(values=wes_palette(n=5, name="Royal2")) +
  labs( x = "Lake Site (By ID #)", y = "Total Obvervations", title = "Top Five sites of Adult and Subadult \n Rana muscosa (Across all years)") +
  theme_light()
  
  
rana_lake_graph

```


Now that have these two graphs of *Rana muscosa*, I am plotting them together, using the patchwork package. 

```{r}
rana_lake_graph / rana_year_graph
```



### End of Document

