---
title: "Task 1"
author: "Dylan Glave"
date: "1/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(patchwork)
library(janitor)
library(here)
library(readxl)
library(gghighlight)
library(ggfortify)
library(plotly)
library(lubridate)
library(ggridges)
library(RColorBrewer)
library(wesanderson)
library(ggtext)
```


## Reading in data for Sierra Nevada Amphibians

```{r}
sierra_raw <- read_excel("sierra_amphibians.xlsx") %>% 
  clean_names() %>% 
  mutate(across(where(is.character), tolower)) %>% 
  mutate(survey_date = ymd(survey_date))
  

```

I've now changed the dates to ymd dates, brought all observations to lower case, and cleaned_names (which didn't seem to do anything.)

Now I'm going to explore visualizing this data. This will require grouping amphibians by year, and creating a df excluding "EggMass," from the life stages **Rana muscosa**

```{r}
ramu_by_year <- sierra_raw %>% 
 mutate(year = year(survey_date)) %>% 
  filter(amphibian_species == "ramu") %>% 
  filter(amphibian_life_stage != "eggmass") 
##Quick summary df of ramu amphibians, grouped by year and life stage.
ramu_summary <- ramu_by_year %>% 
  group_by(year, amphibian_life_stage) %>% 
  summarize(total = sum(amphibian_number, na.rm = TRUE))



```

Making a basic plot of ramu life stages across all sites, grouped by year

```{r}
ggplot(data = ramu_summary, aes(x = year, y = total, fill = amphibian_life_stage)) +
 geom_col() +
  theme_minimal() +
  labs( x = "Year", y ="Total Organisms Found", title =  "Rana muscosa Observations", fill = "Life Stages")

 
```

## Task 1, Part 2: 

We'll play around to get the 5 lakes with the highest counts of *Rana muscosa*

```{r}
ramu_by_lake <-  sierra_raw  %>% 
  filter(amphibian_species == "ramu") %>% 
  filter(amphibian_life_stage != "eggmass") %>% 
  filter(amphibian_life_stage != "tadpole") %>% 
  group_by(lake_id) %>% 
  summarize(total = sum(amphibian_number)) %>% 
  slice_max(total, n=5) %>% 
  mutate(lake_id = as.factor(paste("Lake", lake_id))) %>% 
mutate(lake_id = fct_reorder(lake_id, total))
  
  
```

Now I have a summary table, across all years, of Adult and Subadult Rana Muscosa. In the df, I've made lake_id a factor, added the word "Lake" before each ID number, and reordered the lakes. In their new order, the 5 lakes are listed from highest to lowest counts of *Rana muscosa*

From here, we'll make a column graph of these 5 lakes. 

```{r}
rana_lake_graph <- ggplot(data = ramu_by_lake, aes(x = lake_id, y = total)) +
  geom_col(aes(fill = lake_id, color = lake_id), 
           alpha = 0.8,
           show.legend = FALSE) +
  scale_fill_manual(values=wes_palette(n=5, name="Royal2")) +
  scale_color_manual(values=wes_palette(n=5, name="Royal2")) +
  labs( x = "Lake Site (By ID #)", y = "Total Obvervations", title = "Top Five sites of Adult and Subadult \n Rana muscosa (Across all years)") +
  theme_light()
  
  

rana_lake_graph

```



## Task 2: Principal Component Analysis

I decided to focus this PCA on vegetables. Below I read in the USDA nutrition data, and create a df "veggies_raw" of just Vegetables and Vegetable Products. From there I did some basic cleaning, to make the df coder friendly. 

This Principal Component Analysis 

```{r}
usda_raw <- read.csv("usda_nutrients.csv")

select_foods <- usda_raw %>% 
  filter(FoodGroup == c("Vegetables and Vegetable Products", "Poultry Products", "Snacks")) %>% 
    mutate(across(where(is.character), tolower)) %>% 
  clean_names()

three_food_pca <- select_foods %>% 
  select(energy_kcal,
         ends_with("_g")) %>% 
  drop_na() %>% 
   scale() %>% 
  prcomp()

three_food_pca$rotation

```


```{r}

food_graphable <- select_foods %>% 
drop_na(energy_kcal, ends_with( "_g"))
  

### Then, we make an autoplot of the above PCA, to visualize how Fat and micro nutrients relate to one another. 

autoplot(three_food_pca, 
         data = food_graphable,
         colour = "food_group",
         loadings = TRUE,
         loadings.label = TRUE,
         size = 0.4) +
  theme_minimal() +
  labs( color = "Food Groups") +
  theme_minimal()


```


 - Principal components 1 and 2 cover 76.2% of all variance
-  protein (in grams) is straight up, meaning that is positively correlates with PC2.
 - None of the variables point to the right. This means that none of these variables are positively. 
 - 
