---
title: "Task 1"
author: "Dylan Glave"
date: "1/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(patchwork)
library(janitor)
library(here)
library(readxl)
library(gghighlight)
library(plotly)
library(lubridate)
library(ggridges)
library(RColorBrewer)
```


## Reading in data for Sierra Nevada Amphibians

```{r}
sierra_raw <- read_excel("sierra_amphibians.xlsx") %>% 
  clean_names() %>% 
  mutate(across(where(is.character), tolower)) %>% 
  mutate(survey_date = ymd(survey_date))
  

```

I've now changed the dates to ymd dates, brought all observations to lower case, and cleaned_names (which didn't seem to do anything.)

Now I'm going to explore visualizing this data. This will require grouping amphibians by year, and creating a df excluding "EggMass," from the life stages **Rana muscosa**

```{r}
ramu_by_year <- sierra_raw %>% 
 mutate(year = year(survey_date)) %>% 
  filter(amphibian_species == "ramu") %>% 
  filter(amphibian_life_stage == c("adult", "subadult", "tadpole")) 
##Quick summary df of ramu amphibians, grouped by year and life stage.
ramu_summary <- ramu_by_year %>% 
  group_by(year, amphibian_life_stage) %>% 
  summarize(total = sum(amphibian_number))



```

Making a basic plot of ramu life stages across all sites, grouped by year

```{r}
ggplot(data = ramu_summary, aes(x = year, y = total, fill = amphibian_life_stage)) +
 geom_col() +
  theme_minimal() +
  labs( x = "Year", y ="Total Organisms Found", title =  "Rana muscosa Observations", fill = "Life Stages")

 
```

