---
title: "A2"
author: "Dylan Glave"
date: "1/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(tidyverse)
library(patchwork)
library(janitor)
library(here)
library(readxl)
library(gghighlight)
library(ggfortify)
library(plotly)
library(lubridate)
library(ggridges)
library(RColorBrewer)
library(wesanderson)
library(ggtext)
```

## Task 2: Principal Component Analysis

I decided to focus this PCA on vegetables. Below I read in the USDA nutrition data, and create a df "veggies_raw" of just Vegetables and Vegetable Products. From there I did some basic cleaning, to make the df coder friendly. 

This Principal Component Analysis 

```{r}
usda_raw <- read.csv("usda_nutrients.csv")

select_foods <- usda_raw %>% 
  filter(FoodGroup == c("Vegetables and Vegetable Products", "Poultry Products", "Snacks")) %>% 
    mutate(across(where(is.character), tolower)) %>% 
  clean_names()

three_food_pca <- select_foods %>% 
  select(energy_kcal,
         ends_with("_g")) %>% 
  drop_na() %>% 
   scale() %>% 
  prcomp()

three_food_pca$rotation

```


```{r}

food_graphable <- select_foods %>% 
drop_na(energy_kcal, ends_with( "_g"))
  

### Then, we make an autoplot of the above PCA, to visualize how Fat and micro nutrients relate to one another. 

autoplot(three_food_pca, 
         data = food_graphable,
         colour = "food_group",
         loadings = TRUE,
         loadings.label = TRUE,
         size = 0.4) +
  theme_minimal() +
  labs( color = "Food Groups") +
  theme_minimal()


```

## Some observations on the Biplot:
 - Principal components 1 and 2 cover 76.2% of all variance
- Protein (in grams) is straight up, meaning that is positively correlates with PC2.
 - None of the variables point to the right. This means that none of these variables are positively. 
 - Variables, shown in red lines, that are close in angle, are very correlated.
 - The same goes for points! Points that appear in clusters are closely related across the first principal components. In this plot, the *Poultry* and *Vegetables* food groups form clusters, while very few of the *Snack* points cluster together.
 
